A* with Perfect Potentials

### 2020-06-09 paper/ch_potentials.tex
foo

To prove correctness, consider the turn-aware path weight of an optimal path $v_1,v_2\ldots v_k$ with $v_1=x$, $v_2=y$, $v_{k-1}=p$, and $v_k=q$.
We can lower bound the expression as follows:
\[
\sum_{i=2}^{k-1} w_\ell(v_i,v_{i+1}) \le \sum_{i=2}^{k-1} \underbrace{w_t(v_{i-1},v_i,v_{i+1})}_{0\le} + \sum_{i=2}^{k-1} \underbrace{w_q(v_i,v_{i+1})}_{w_\ell(v_i,v_{i+1})\le}
\]
The $\sum_{i=2}^{k-1} w_\ell(v_i,v_{i+1})$ expression is the length of some, not necessarly shortest, $yq$-path in $G$ with $w_\ell$.
It is by definition not smaller than the shortest $yq$-path length in $G$ with $w_\ell$.
The shortest $yq$-path length is exactly what CH-Potentials compute.
The potential is therefore a lower bound.
It is consistent because it is derived from exact shortest path in $G$ with $w_\ell$.

@@

With this setup we extended CH-Potentials in a way that keeps all turn-related information out of the CH and the preprocessing.
This allows for a comparatively easy setup and turn-aware routing can be activate or deactivate for every query individually.

\subsection{Time-Dependent Routing}

>>>>

To prove correctness, consider the turn-aware path weight of an optimal path $v_1,v_2\ldots v_k$ with $v_1=x$, $v_2=y$, $v_{k-1}=p$, and $v_k=q$.
We can lower bound the expression as follows:
\[
\sum_{i=2}^{k-1} w_\ell(v_i,v_{i+1}) \le \sum_{i=2}^{k-1} \underbrace{w_t(v_{i-1},v_i,v_{i+1})}_{0\le} + \sum_{i=2}^{k-1} \underbrace{w_q(v_i,v_{i+1})}_{w_\ell(v_i,v_{i+1})\le}
\]
The $\sum_{i=2}^{k-1} w_\ell(v_i,v_{i+1})$ expression is the length of some, not necessarly shortest, $yq$-path in $G$ with $w_\ell$.
It is by definition not smaller than the shortest $yq$-path length in $G$ with $w_\ell$.
The shortest $yq$-path length is exactly what CH-Potentials compute.
The potential is therefore a lower bound.
It is consistent because it is derived from exact shortest paths in $G$ with $w_\ell$.

@@

With this setup we extended CH-Potentials in a way that keeps all turn-related information out of the CH and its preprocessing.
This allows for a comparatively easy setup and turn-aware routing can be activate or deactivate for every query individually.

\subsection{Historic Traffic Patterns}

The classical shortest path problem assumes that edge weights are scalars. 
However in practice, travel times vary along an edge.
One source of variance are recurring congestion patterns.

These are modelled using functions as edge weights~\cite{lots-of-td-papers}.
They map the time of entry into an edge onto the travel time through it.
Formally, $w_q$ is now a \emph{travel time function} from $E\times \mathbb{R}$ to $\mathbb{R}^+$. 
$w_q(e, \tau)$ is the travel time through edge $e$ when entering it at $\tau$.

The input to the problem consists of a source node $s$ and a target node $t$ as in the classical problem formulation.
Additionally, the input contains a source time $\tau_s$.
A path with edges $e_1,e_2\ldots e_k$ is weighted using $\alpha_k$, which is defined recursively as follows:\[
\begin{split}
\alpha_{1} & = w_q(e_1, \tau_s) \\
\alpha_{k} & = \alpha_{k-1} + w_q(e_1, \alpha_{k-1})
\end{split}
\]
The objective is to find a path that minimizes $\alpha_k$.

If all travel time functions fulfill the \emph{no-waiting property}, this problem can be solved using a straight forward extension of Dijkstra's algorithm \cite{td-alt?}.
The necessary modification to A* is analogous.
Without the no-waiting property the problem is in general NP-hard \cite{veit}.
The no-waiting property states that it is never benficial to wait at a node before entering an edge.
Formally stated, the following must hold:\[
\forall e\in E,\tau\in \mathbb{R},\delta\in \mathbb{R}^+: w_q(e, \tau) \le w_q(e, \tau+\delta) + \delta
\]

Usually, travel time functions are stored as piece-wise linear functions.
Our implementation follows this setup.
However in priciple, our algorithm works with any representation that enables an efficient function evaluation.

Following \cite{td-alt}, we set $w_\ell(e)$ to the minimum travel time for every edge, i.e., $w_\ell(e) = \min_\tau w_q(e,\tau)$.
This setup gives us a lower bound weight that is independent of the edge entry time. 
Howeve, for every entry time, $w_\ell(e)$ is a valid lower bound.

The modifications to make CH-Potentials work are straight forward.
We compute the CH with respect to $w_\ell$ which maps edges onto scalars.
With this CH, we can compute CH-Potentials.
These are then combined with the travel-time-function-aware A* variant of \cite{td-alt?}.

With this setup we extended CH-Potentials in a way that keeps all travel-time-function-related information out of the CH and its preprocessing.
This way, we avoid the main problems of \cite{tdch,catchup}, which propagate the travel time functions into the CH construction.

\subsection{Live Traffic}
